Abstract Data Type Inference

### 2015-03-18 patterns.tex
Added a figure.
started wqo proof

\section{Computing ADT Complements}
\label{sec:patterns}

@@

@@

>>>>

%!TEX root = draft.tex
\section{Computing ADT Complements}
\label{sec:patterns}

@@

\begin{proof}

%We prove that $\preceq$ is well-founded, i.e., there is no infinite strictly decreasing sequence $h_1 \succ h_2\succ \ldots$\footnote{We define $\succ$ as usual by $h\succ h'$ iff $h\succeq h'$ and $h\neq h'$.},
%and that $\preceq$ has no infinite antichain, i.e., there is no infinite set $\{h_1,h_2,\ldots\}$ of mutually incomparable elements 
%$h_i\not\preceq h_j$ and $h_j\not\preceq h_i$ when $i\neq j$.
%
%Concerning well-foundedness, note that a given history $h_1$ can be weaken only finitely many times by removing matches, duplicates in a match, or happens-before ordering constraints, or by turning completed operations into pending operations. Since we restrict ourselves to width-bounded histories, $h_1$ can also be weaken only finitely many times by adding pending operations (adding a pending operation increases the width of a history by 1).
%
%TODO PROBLEM WITH PENDING OPS

%To prove that $\preceq$ has no infinite antichain, 
%we use the following definitions.
Let $h_1 h_2\ldots $ be an infinite sequence of histories. We prove that there exists $i<j$ such that $h_i\preceq h_j$.

The \emph{signature} of an operation $o$ in a history $h$ is the pair $\sigma(o)=(c(o),f(o))$.
Then, the \emph{signature} of a match $\mu=\set{ o' \in O : m(o') = o}$ is the pair 
\[
\sigma(\mu)=(\sigma(o),\set{\sigma(o') : m(o')=o, o'\neq o})
\]
containing the signature of the match target and the set of the signatures of the other operations.
The \emph{signature} of a history $h$ is the set 
\[
\sigma(h)=\set{\sigma(\mu): \mu\mbox{ a match of $h$}}\cup\set{\sigma(o): o\mbox{ unmatched}}.
\]
of match signatures and signatures of unmatched operations.

%Assume by contradiction that $\preceq$ has an infinite antichain.
Since the set of history signatures is bounded (because the set of methods is bounded)
and the set of pending operations is also bounded (since we consider only bounded width histories),
the sequence $h_1 h_2\ldots $ contains an infinite sequence of histories $h_1' h_2' \ldots$ that have 
the same signature and the same number of pending operations with the same signature.

%The \emph{vector} of a match $\mu$ is the pair
%\[
%\nu(\mu)=(\sigma(o),\mset{\sigma(o') : m(o')=o, o'\neq o})
%\]
%containing the signature of the match target and the multiset of the signatures of the other operations.
%Also, 
The \emph{vector} of a history $h$ is the multiset 
\[
\nu(h)=\mset{\sigma(\mu): \mu\mbox{ a match of $h$}}
\]
of match signatures in $h$.

Given a finite alphabet $\Sigma$, a domain $\mathbb{D}$ enhanced with a wqo $\leq$, and two functions 
$\alpha,\beta:\Sigma\rightarrow \mathbb{D}$, let $\sqsubseteq$ be an order relation defined by
\[
\alpha\sqsubseteq\beta\mbox{ iff for every $a\in \Sigma$, $\alpha(a)\leq\beta(a)$}.
\]
Note that $\sqsubseteq$ is a wqo.

The relation $\sqsubseteq$ on history vectors (mappings from the alphabet $\Sigma$ of match signatures to $\mathbb{N}$), 
where $\leq$ is the order relation on natural numbers, is a wqo.
Hence, the set $H$ contains an infinite sequence $h_1',h_2',\ldots$ such that $\nu(h_1')\sqsubseteq \nu(h_2')\sqsubseteq \ldots$.

A \emph{nested vector} of a history $h$ is the nested multiset
\[
\nu^2(h)=\mset{\mset{\nu(\mu): \sigma(\mu)=\sigma} : \sigma\mbox{ the signature of a match in $h$}}.
\]

Again using the fact that $\leq$ is a wqo, the sequence $h_1',h_2',\ldots$ contains a history $h'_k$ and an infinite set of histories 
$\set{h_1'',h_2'',\ldots}$ such that for every $i$ and every match $\mu$ of $h_i''$,
$\nu(\mu_k)\leq \nu(\mu)$, for some match $\mu_k$ of $h_k$ such that $\sigma(\mu_k)=\sigma(\mu)$.

\end{proof}

@@

  \begin{figure}[t]
    \centering
    \includegraphics[scale=0.2]{figures/register-pattern-1},
    \includegraphics[scale=0.2]{figures/register-pattern-2},
    \includegraphics[scale=0.2]{figures/register-pattern-3}
    \caption{Three histories which generate the complement of the atomic
      register ADT kernel.}
    \label{fig:register-patterns}
  \end{figure}
